Equation-of-state insensitive relations after GW170817

### 2019-04-10 paper/Response.txt

Zack Carson, Katerina Chatziioannou, Carl-Johan Haster, Kent Yagi, and 
Nicolás Yunes

---------------------------------------
Comments from the Referee
---------------------------------------

R: 1. I find the description of the “constrained” prior and its relation
to the “unconstrained” one slightly confusing. Sec. I A suggests that
it is constructed by sampling a large number of spectral EOSs and then
discarding those lying outside the 90% credible region of the EOS
posterior from GW170817. On the other hand, Sec. II A says that it is
built up from EOSs sampled directly from the EOS posterior (itself
derived from the unconstrained prior?), with no mention of the 90%
credible region. Naively, I would think that one would simply take the
unconstrained prior and reweight it by the GW170817 EOS likelihood to
get the constrained prior. Please explain more clearly which procedure
was adopted, and emphasize how the constrained and unconstrained
priors are related.

A: We apologize for the loss in translation between the descriptions of 
the constrained EoSs in the executive summary and Sec IIa. Ultimately, 
the "constrained" EoSs were computed identically to how the
posterior for the EoS was computed in the paper by Abbott et al. So 
essentially we are sampling the actual posterior given by LIGO (the exact 
posterior EoSs from the paper were not publicly available at the time we 
started this work, so the computation had to be re-done). Then, the 
unconstrained EoSs were generated completely randomly, with no relation 
to the constrained EoS or the posterior PDF. We have modified our 
descriptions of this in both Sec. Ia and IIa.

@@

R: 2. To quantify the error in the tidal deformability introduced by the
approximately universal nature of the binary Love relations, Sec. IV
adopts a representation for the error about the best-fit relation. The
error model is fitted to data from the constrained prior. Does this
prior include any hybrid EOSs, the models that deviate the most from
the binary Love relations? If not, do the projections in Sec. IV rely
on an implicit assumption that the EOS is purely hadronic?

A: Yes, you are correct that our results for the systematic errors from
the binary Love relations rely on the assumption of hadronic stars. Our 
set of constrained EoSs do not contain any hybrid star EoSs, following
from the calculations of Abbott et al on the EoS posterior probability 
distribution, which also assumed purely hadronic EoSs. If included these
results may change slightly, and could be a possible avenue of future
work. We have added a couple of sentences to the end of Sec. IVa 
noting this.

R:  3. Several papers have claimed that EOSs with sharp phase transitions
can violate I-Love-Q universality, e.g. Han & Steiner,
arXiv:1810.10967; Bandyopadhyay et al., Eur. Phys. J. A (2018) 54: 26;
Lau et al., Phys. Rev. D 95, 101302 (2017). Here, in Sec. III, hybrid
hadron-quark EOSs are found to obey the standard I-Love-Q relations.
Is this finding compatible with the existing claims in the literature?
The paper would benefit from a brief discussion of this point.

A: Upon reading over the papers by Han & Steiner, Bandyopadhyay et al, 
and Lau et al, we find that our results are indeed consistent with what
they present. 
The paper by Han & Steiner claims that "for typical neutron
stars > 1.0 M_sun, discrepancies are negligible", which can also be seen
in the bottom left panel of Fig. 11. In fact, the fits are only valid in
the M>1M_sun regime as can be seen in the top panel of Fig. 11, a relic 
fixed in our paper by using improved fits which limit to the Newtonian
case.
In the paper by Lau et al,  they claim that the "I-Love 
relation for hybrid stars still agrees with the universal relation for 
fluid stars to about 1% level for hybrid stars with a quark-matter core
surrounded by a fluid [hadronic] envelope", similar to what we investigate. 
They find that solid quark-matter stars (not studied by us) violate the 
relations proportionally to the value of the gap parameter. 
Finally, in the paper by Bandyopadhyay et al, there is no claim to *how* 
strongly the "HQ" EoSs violate the I-Love relation. However, by looking 
at Fig. 6 we see that the departure from the "np" EoSs is not noticeably 
large, and seems to be somewhat consistent with what we see with our 
hybrid star EoSs. 
We also see in Paschalidis et al. 2018 (10.1103/physrevd.97.084038) that it was 
shown that the I-Love-Q universality is preserved for the hybrid stars
considered, which are identical to the ones used here.
As so, we have added a couple of sentences at the end of Sec.
IIIA2 discussing the consistency, as well as the additional references.

### 2019-04-10 paper/Submit/bin-love-after-gw170817.tex
implementing Katerina and Kents comments on response to Ref.

%How we improved this
We begin by investigating the increase in EoS insensitivity due to the constraints placed by GW170817 on the allowed space of EoSs. 
We take the 90\% credible region constructed from the posterior probability distributions on the pressure-density plane~\cite{Carney:2018sdv,LIGO:posterior} obtained from GW170817~\cite{TheLIGOScientific:2017qsa} as a prior on the set of EoSs that is compatible with this observation. 
We then generate two large samples of spectral EoSs~\cite{Lindblom:2018rfr}, one in which \textbf{the EoSs are directly sampled from the posterior probability distribution} (the ``constrained'' sample) and another in which the prior is not enforced (the ``unconstrained'' sample). 
We repeat the analysis done by Yagi and Yunes~\cite{Yagi:binLove,Yagi:ILQ} on both sets of EoSs and find that the EoS-insensitive relations present less EoS variability with the constrained set. 
In particular, the EoS-insensitivity increases by a factor of $\sim 60$\% in the binary Love relations (for stars with mass ratio larger than $0.75$), by a factor of $\sim 70\%$ in the C-Love relations, and by factors of $\sim 50$\% in the I-Love-Q relations (see Table~\ref{tab:maxVar} and Fig.~\ref{fig:ILQ} for more details).

@@

The unconstrained set is then defined by drawing random samples in the spectral coefficients within their allowed prior ranges, and then eliminating any EoS that either leads to \textbf{an adiabatic index} $\Gamma$ outside the allowed range, breaks the causality restriction, or breaks the maximum mass restriction. 
To obtain the constrained set of EoSs we analyze the publicly available data for GW170817 from the Gravitational Wave Open Science Center~\cite{GWOSC,Vallisneri:2014vxa}. 
We use the publicly available software library {\tt LALInference}~\cite{Veitch:2014wba,lalinference_o2} to \textbf{directly} obtain samples from the EoS \textbf{90\% posterior provided by Ref~\cite{LIGO:posterior}.}
Our analysis uses the same settings and prior choice as those of~\cite{LIGO:posterior,Carney:2018sdv}. 
In both cases, each set consists of 100 members, with a subset of these shown in the left panel of Fig.~\ref{fig:eos}.

>>>>

%How we improved this
We begin by investigating the increase in EoS insensitivity due to the constraints placed by GW170817 on the allowed space of EoSs. 
\textbf{We use the posterior probability distribution on the pressure-density plane~\cite{Carney:2018sdv,LIGO:posterior} obtained from GW170817~\cite{TheLIGOScientific:2017qsa} to inform the set of EoSs that is compatible with this observation. We then generate two large samples of spectral EoSs~\cite{Lindblom:2018rfr}, one in which the EoSs are directly sampled from the posterior probability distribution (the ``constrained" sample) and another in which this constraint not enforced (the ``unconstrained" sample).}
We repeat the analysis done by Yagi and Yunes~\cite{Yagi:binLove,Yagi:ILQ} on both sets of EoSs and find that the EoS-insensitive relations present less EoS variability with the constrained set. 
In particular, the EoS-insensitivity increases by a factor of $\sim 60$\% in the binary Love relations (for stars with mass ratio larger than $0.75$), by a factor of $\sim 70\%$ in the C-Love relations, and by factors of $\sim 50$\% in the I-Love-Q relations (see Table~\ref{tab:maxVar} and Fig.~\ref{fig:ILQ} for more details).

@@

The unconstrained set is then defined by drawing random samples in the spectral coefficients within their allowed prior ranges, and then eliminating any EoS that either leads to \textbf{an adiabatic index} $\Gamma$ outside the allowed range, breaks the causality restriction, or breaks the maximum mass restriction. 
To obtain the constrained set of EoSs we analyze the publicly available data for GW170817 from the Gravitational Wave Open Science Center~\cite{GWOSC,Vallisneri:2014vxa}. 
We use the publicly available software library {\tt LALInference}~\cite{Veitch:2014wba,lalinference_o2} to sample the EoS \textbf{posterior, similar to Ref~\cite{LIGO:posterior}.}
Our analysis uses the same settings and prior choice as those of~\cite{LIGO:posterior,Carney:2018sdv}. 
In both cases, each set consists of 100 members, with a subset of these shown in the left panel of Fig.~\ref{fig:eos}.
